Fully Automatic Loop-free Program Synthesis by Bounded Model Checking

### 2014-01-31 paper/paper.tex
Typo
Added benchmarks appendix

\usepackage{listings}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{pgf}
\usepackage[noend]{algpseudocode}
\usepackage{algorithm}
\usepackage{amsfonts}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{pifont}
\usepackage{multicol}

@@

\section{Experiments}
\label{sec:experiments}
\subsection{Experimental Setup}
We implemented our fully automatic synthesis procedure as the {\sc kalashnikov} tool.  The implementation consists
of around 600 lines of \newC and 600 lines of Python.  To evaluate the tool we used the 25 bitvector
programs from~\cite{brahma} and~\cite{brahma-icse}.  The majority of these are ``bit twiddling hacks'' taken from
Hacker's Delight~\cite{hackers-delight}.  The code we used to perform the experiments, along with the
benchmarks, is available at \texttt{http://www.cprover.org/kalashnikov}.  As a backend solver, we used
{\sc CBMC}~\cite{cbmc-website} at SVN revision r3545, with Glucose 3.0~\cite{glucose-paper} as the SAT solver.
We performed our experiments on a 4-core, 2.40\,GHz Xeon E5-2665 with 32\,GB of RAM.

@@

\paragraph{Acknowledgements}
We thank Martin Brain for his insight and advice during our synthesis discussions.
Also, thanks to Cristina David for her extensive help and proof reading during the
preparation of this paper.

@@

>>>>

\usepackage{listings}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{pgf}
\usepackage[noend]{algpseudocode}
\usepackage{algorithm}
\usepackage{amsfonts}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{pifont}
\usepackage{multicol}
\usepackage{appendix}

@@

\section{Experiments}
\label{sec:experiments}
\subsection{Experimental Setup}
We implemented our fully automatic synthesis procedure as the {\sc kalashnikov} tool.  The implementation consists
of around 600 lines of \newC and 600 lines of Python.  To evaluate the tool we used the 29 bitvector
programs from~\cite{brahma} and~\cite{brahma-icse}.  The majority of these are ``bit twiddling hacks'' taken from
Hacker's Delight~\cite{hackers-delight}.  The code we used to perform the experiments, along with the
benchmarks, is available at \texttt{http://www.cprover.org/kalashnikov} and the benchmarks can also
be found in Appendix~\ref{app:benchmarks}.  As a backend solver, we used
{\sc CBMC}~\cite{cbmc-website} at SVN revision r3545, with Glucose 3.0~\cite{glucose-paper} as the SAT solver.
We performed our experiments on a 4-core, 2.40\,GHz Xeon E5-2665 with 32\,GB of RAM.

@@

\paragraph{Acknowledgements}
We thank Martin Brain for his insight and advice during our synthesis discussions.
Also, thanks to Cristina David for her extensive help and proofreading during the
preparation of this paper.

@@

\appendix

\section{Benchmarks}
\label{app:benchmarks}

\begin{figure}
\scriptsize
 \begin{multicols}{4}

   \begin{minipage}{.2\textwidth}
   \bf{P1}($x$): Turn-off rightmost 1 bit.
   \begin{enumerate}
    \item $t1 = x - 1$
    \item $t2 = x \& t1$
   \end{enumerate}
   \end{minipage}

   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   \bf{P2}($x$): Test whether $x$ is of the form $2^n - 1$.
   \begin{enumerate}
    \item $t1 = x + 1$
    \item $t2 = x \& t1$
   \end{enumerate}
   \end{minipage}

   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   \bf{P3}($x$): Isolate the rightmost 1 bit.
   \begin{enumerate}
    \item $t1 = -x$
    \item $t2 = x \& t1$
   \end{enumerate}
   \end{minipage}

   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   \bf{P4}($x$): Form a mask that identifies the rightmost 1 bit and trailing 0s.
   \begin{enumerate}
    \item $t1 = x - 1$
    \item $t2 = x \oplus t1$
   \end{enumerate}
   \end{minipage}

   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   \bf{P5}($x$): Right propagate the rightmost 1 bit.
   \begin{enumerate}
    \item $t1 = x - 1$
    \item $t2 = x \mid t1$
   \end{enumerate}
   \end{minipage}

   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   \bf{P6}($x$): Turn on the rightmost 0 bit.
   \begin{enumerate}
    \item $t1 = x + 1$
    \item $t2 = x \mid t1$
   \end{enumerate}
   \end{minipage}
      \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   \bf{P7}($x$): Isolate the rightmost 0 bit.
   \begin{enumerate}
    \item $t1 = \sim x$
    \item $t2 = x + 1$
    \item $t3 = t1 \& t2$
   \end{enumerate}
   \end{minipage}
   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   \bf{P8}($x$): Form a mask the identifies the trailing 0s.
   \begin{enumerate}
    \item $t1 = x -1$
    \item $t2 = \sim x$
    \item $t3 = t1 \& t2$
   \end{enumerate}
   \end{minipage}
   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   \bf{P9}($x$): Absolute value.
   \begin{enumerate}
    \item $t1 = x >> 31$
    \item $t2 = x \oplus t1$
    \item $t3 = t2 - t1$
   \end{enumerate}
   \end{minipage}

      \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   \bf{P10}($x, y$): Test if nlz(x) == nlz(y), where nlz is number of leading zeroes.
   \begin{enumerate}
    \item $t1 = x \& y$
    \item $t2 = x \oplus y$
    \item $t3 = t2 \leq t1$
   \end{enumerate}
   \end{minipage}

      \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   \bf{P11}($x, y$): Test if nlz(x) < nlz(y).
   \begin{enumerate}
    \item $t1 = \sim y$
    \item $t2 = x \& t1$
    \item $t3 = t2 > y$
   \end{enumerate}
   \end{minipage}
   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   \bf{P12}($x, y$): Test if nlz(x) <= nlz(y).
   \begin{enumerate}
    \item $t1 = \sim y$
    \item $t2 = x \& t1$
    \item $t3 = t2 \leq y$
   \end{enumerate}
   \end{minipage}

      \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   \bf{P13}($x$): Sign.
   \begin{enumerate}
    \item $t1 = x >> 31$
    \item $t2 = t1 < x$
    \item $t3 = t2 + t1$
   \end{enumerate}
   \end{minipage}

   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   \bf{P14}($x, y$): Floor of average of $x$ and $y$ without overflowing.
   \begin{enumerate}
    \item $t1 = x \& y$
    \item $t2 = x \oplus y$
    \item $t3 = t2 >> 1$
    \item $t4 = t1 + t3$
   \end{enumerate}
   \end{minipage}
   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   \bf{P15}($x, y$): Ceil of average of $x$ and $y$ without overflowing.
   \begin{enumerate}
    \item $t1 = x \mid y$
    \item $t2 = x \oplus y$
    \item $t3 = t2 >> 1$
    \item $t4 = t1 - t3$
   \end{enumerate}
   \end{minipage}
   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   \bf{P16}($x$): Max.
   \begin{enumerate}
    \item $t1 = x \leq y$
    \item $t2 = x \oplus y$
    \item $t3 = t1 * t2$
    \item $t4 = t3 \oplus y$
   \end{enumerate}
   \end{minipage}
   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   \bf{P17}($x$): Turn off the rightmost contiguous string of 1 bits.
   \begin{enumerate}
    \item $t1 = x - 1$
    \item $t2 = x \mid t1$
    \item $t3 = t2 + 1$
    \item $t4 = t3 \& x$
   \end{enumerate}
   \end{minipage}
   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   \bf{P18}($x$): Determine if $x$ is a power of 2.
   \begin{enumerate}
    \item $t1 = -x$
    \item $t2 = x \oplus t1$
    \item $t3 = t2 < t1$
   \end{enumerate}
   \end{minipage}
   \vspace{12pt}

  \iffalse
   \begin{minipage}{.2\textwidth}
   \bf{P19}($x, m, k$): Exchange two fields of $x$, demarcated by $m$ and $k$.
   \begin{enumerate}
    \item $t1 = x >> k$
    \item $t2 = x \oplus t1$
    \item $t3 = t2 & m$
    \item $t4 = t3 << k$
    \item $t5 = t4 \oplus t3$
    \item $t6 = t5 \oplus x$
   \end{enumerate}
   \end{minipage}
   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   \bf{P20}($x$): Next higher unsigned number with the same number of 1 bits.
   \begin{enumerate}
    \item $t1 = \sim x$
    \item $t2 = x \& t1$
    \item $t3 = x + t2$
    \item $t4 = x \opus t2$
    \item $t5 = t4 >> 2$
    \item $t6 = t5 / t2$
    \item $t7 = t6 \mid t3$
   \end{enumerate}
   \end{minipage}
   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   \bf{P21}($x, a, b, c$): Cycling through 3 values $a, b, c$.
   \begin{enumerate}
    \item $t1 = x \neq c$
    \item $t2 = a \oplus c$
    \item $t3 = x \neq a$
    \item $t4 = b \oplus c$
    \item $t5 = t2 \& t2$
    \item $t6 = t3 \& t4$
    \item $t7 = t5 \oplus t6$
    \item $t8 = t7 \oplus c$
   \end{enumerate}
   \end{minipage}
\fi

   \begin{minipage}{.2\textwidth}
   {\bf P19($x, m, k$): Exchange two fields of $x$, demarcated by $m$ and $k$.}

   N/A ({\sc Kalashnikov} was unable to synthesise this program.)
   \end{minipage}
   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   {\bf P20($x$): Next higher unsigned number with the same number of 1 bits.}

   N/A
   \end{minipage}
   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   {\bf P21($x, a, b, c$): Cycling through 3 values $a, b, c$.}

    N/A
   \end{minipage}
   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   {\bf P22($x$): Compute parity.}

    N/A
   \end{minipage}
   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   {\bf P23($x$): Count number of set bits.}

    N/A
   \end{minipage}
      \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   {\bf P24($x$): Round up to next highest power of 2.}

    N/A
   \end{minipage}
   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   {\bf P25($x, y$): Computer higher order half of $x \times y$.}

    N/A
   \end{minipage}

   \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   {\bf P26($x, k$): Round $x$ up to a multiple of $2^k$.}

N/A
   \end{minipage}

      \vspace{12pt}

   \begin{minipage}{.2\textwidth}
   {\bf P27($x, y$): Min.}
    \begin{enumerate}
     \item $t1 = x \oplus y$
     \item $t2 = t1 \leq t2$
     \item $t3 = t1 * t2$
     \item $t4 = y \oplus t3$
    \end{enumerate}
   \end{minipage}

 \end{multicols}

 \caption{The integer benchmarks}
\end{figure}

\begin{figure}
 \begin{lstlisting}[language=C++]
void interchangeObs(word_t* src, word_t* dest)
{
  *src = *src ^ *dest;

  if (*src == (*src ^ *dest)) {
    *src = *src ^ *dest;

    if (*src == (*src ^ *dest)) {
      *dest = *src ^ *dest;

      if (*dest == (*src ^ *dest)) {
        *src = *dest ^ *src;
        return;
      } else {
        *src = *src ^ *dest;
        *dest = *src ^ *dest;
        return;
      } 
    } else
      *src = *src ^ *dest;
  }

  *dest = *src ^ *dest;
  *src = *src ^ *dest;

  return;
}
 \end{lstlisting}

 \caption{Benchmark 28: an obfuscated swap function}
\end{figure}

\begin{figure}
 \begin{lstlisting}[language=C++]
int mul45(int y) {
  int a=1, b=0, z=1, c=0;

  while(1) {
    if (a == 0) {
      if (b == 0) {
	y=z+y; a =!a; b=!b;c=!c;
	if (! c) break;
      } else {
	z=z+y; a=!a; b=!b; c=!c;
	if (! c) break;
      }
    } else {
      if (b == 0) { z=y << 2; a=!a; }
      else {
	z=y << 3; a=!a; b=!b; } } }

  return y;
}

 \end{lstlisting}

 \caption{Benchmark 29: an obfuscated function to multiply by 45}
